A Possibilistic Valid-time Model

### 2011-12-01 trunk/ipmu2012/preliminaries.tex
More changes, resulting in one less page...

\subsection{\label{subsec:possibilistic-variables}Possibilistic Variables}
A \emph{possibilistic variable} can intuitively be seen as a variable for which the possible values are described by its possibility distribution\cite{Pon11}.

\begin{definition}
A possibilistic variable $X$ over a universe $U$ is defined as a variable taking exactly one value in $U$, but for which this value is (partially) unknown. The possibility distribution $\pi_X$ on $U$ gives the available knowledge about the value that $X$ takes: for each $u\in U$, $\pi_X(u)$ represents the possibility that $X$ takes the value $u$.
\end{definition}

@@

When a possibilistic variable is defined on the powerset $\Pow(R)$ of some universe $R$, the unique value the variable takes will be a crisp set. In this case, the exact set is off course (partially) unknown, but the available knowledge is given in the form of a possibility distribution on the powerset $\Pow(R)$, which describes the possibility of each crisp subset of $R$ to be the value the variable takes. This exact value (a crisp set) the variable takes, is now called an \emph{ill-known set}\cite{Dubois88b}.

@@

A specific application of possibilistic variables is obtained when the universe under consideration is the set of boolean values $\mathbb{B}$ = $\{T,F\}$. Indeed, any boolean proposition $p$ takes just one value in $\mathbb{B}$. If the knowledge on which value this proposition $p$ will take is given by a possibility distribution $\pi_p$, the proposition can be seen as a possibilistic variable. As the interest lies with the case where the proposition holds, the possibility and necessity that $p$ = $T$ (the proposition holds) demands most attention. This possibility and necessity is noted here as:
\begin{align}
Pos(p) & = \pi_p(T) \label{propholdsposs} \\
Nec(p) & = 1-\pi_p(F) \label{propholdsnecc}
\end{align}
Here, notation \ref{propholdsposs} denotes the possibility that $p$ = $T$ and the proposition holds, notation \ref{propholdsnecc} denotes the necessity that $p$ = $T$ and the proposition holds.

\subsection{\label{subsec:fuzzy-numbers}Fuzzy Numbers and Fuzzy Intervals}
Among others, Dubois and Prade~\cite{Dubois1983} use the following definition of a \emph{fuzzy interval}.
\begin{definition}
A fuzzy interval is a fuzzy set $M$, defined by a membership function $\mu_{M}$, on the set of real numbers $\mathbb{R}$ such that:
\begin{eqnarray}
\mu_{M} : & \!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\! \mathbb{R} \rightarrow \left[0,1\right] \nonumber \\ 
\forall (u,v)\in\mathbb{R}^2: \forall w \in [u,v]:&\mu_M(w) \geq\min(\mu_M(u),\mu_M(v))  \\
\exists m \in \mathbb{R} : & \!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\! \mu_M(m)=1 
\end{eqnarray}
\end{definition}
If this modal value $m$ is unique, then $M$ is referred to as a \emph{fuzzy number}, instead of a \emph{fuzzy interval}. In other words, if the core of a fuzzy interval is a singleton, it is referred to as a fuzzy number.

A simple form of the membership function of a fuzzy interval is a trapezoidal function. It can be shown that such a membership function $\mu_T$ for a fuzzy interval $T$ is convex and normalized. A visualization is given in figure \ref{fig:trapezoidal}. Four reel values, denoted $\alpha$, $\beta$, $\gamma$ and $\delta$ and chosen as in figure \ref{fig:trapezoidal}, suffice to represent a trapezoidal membership function of a fuzzy interval. In this work, a fuzzy interval defined as such will be noted as $\left[\alpha, \beta, \gamma, \delta\right]$. The corresponding membership function definition for this $\mu_T$ is then given by:

\begin{align}
\mu_T : & \quad \mathbb{R} \rightarrow \left[0,1\right] \\
 : & \quad x \rightarrow
\begin{cases}
1 & \mbox{ if } x \in [\beta,\gamma] \\
0 & \mbox{ if } x > \delta \vee x < \alpha \\
\frac{x-\alpha}{\beta - \alpha} & \mbox{ if } x \in [\alpha,\beta[ \\
\frac{\delta -x}{\delta - \gamma} & \mbox{ if } x \in ]\gamma,\delta] \\
\end{cases}
\end{align}

@@

The most convenient form of the membership function of a fuzzy number is a triangular function. It can be shown that such a membership function $\mu_M$ for a fuzzy number $M$ is convex and normalized. A visualization is given in figure \ref{fig:triangular}. Three reel values, denoted $a$, $b$ and $D$, suffice to represent a triangular membership function of a fuzzy number.
\begin{itemize}
\item
$D$ denotes the single value in the core of $M$
\item
$D-a$ is then $\inf \{u \in \mathbb{R} : \mu_{M}(u) > 0\}$
\item
$D+b$ is then $\sup \{u \in \mathbb{R} : \mu_{M}(u) > 0\}$
\end{itemize}
In this work, a fuzzy number defined as such will be noted as $\left[D, a, b \right]$.
\begin{figure}[h!]
  \centering
  \input{./graphs/triangular.tex}
  \caption{Triangular membership function.}
  \label{fig:triangular}
\end{figure}

@@

\subsubsection{Interval Evaluation by Ill-known Constraints}
The problem of interval evaluation is more generally explained in \cite{Pon11}: for a crisp interval $I = \left[a, b \right]$, the need exists to know if all points in this interval reside between the boundaries of an ill-known interval $\left[ X , Y \right]$.

The notion of an \emph{ill-known constraint} is introduced:

\begin{definition}
Given a universe $U$, an \emph{ill-known constraint} $C$ on a set $A \subseteq U$ is specified by means of a binary relation $R \subseteq U^{2}$ and a fixed, ill-known value denoted by its variable $X$ over $U$, i.e.:
\begin{align}
C \triangleq (R,X)
\end{align}
\end{definition}

We assume that $X$ specifies the lower bound and $Y$ the upper bound for a given interval, we want to known whether all points in the interval are larger than or equal to $X$ and smaller than or equal to $Y$. Therefore, we consider two ill-known constraints:

\begin{eqnarray}
C_1\stackrel{\triangle}{=}\left(\leq,X\right)\\
C_2\stackrel{\triangle}{=}\left(\geq,Y\right).
\end{eqnarray}

The following expressions compute the possibility and the necessity:

\begin{eqnarray}
\label{eq:interval-pos}
\Pos\left(\lambda([a,b])\right)=\\
\nonumber
\min\bigg(\sup_{a\geq w}\pi_{X}(w),\sup_{b\leq w}\pi_{Y}(w)\bigg)\\
\label{eq:interval-nec}
\Nec\left(\lambda([a,b])\right)=\\
\nonumber
\min\bigg(\inf_{a<w}1-\pi_{X}(w),\inf_{b>w}1-\pi_{Y}(w)\bigg).
\end{eqnarray}

@@

\subsection{Temporal Databases}
%In this section, the proposed reasoning is applied to the specific context of intervals on the real line. This setting is of specific interest in the context of fuzzy temporal databases.
A temporal database \cite{Dyreson1994} is a database that manages some aspects of time in its schema. A \textbf{chronon} is the shortest duration of time supported by the database. The time can be represented either as points or intervals \cite{655777}. Fuzzy temporal models \cite{schockaert08} are proposed when the time points or intervals are not precisely known. There are proposals for the  fuzzyfication of the time point and the fuzzyfication of the time interval.\\
Allen \cite{Allen83} studied the comparison between two crisp time intervals. For fuzzy intervals, several proposals \cite{ohlbach04,nagypal03,schockaert08} have been done.

@@

Despite of user-defined time (an uninterpreted attribute; supported in the standard SQL2), there are 3 types of time in a temporal database:

\begin{itemize}
	\item
	\textbf{Transaction time}: The time when the fact is stored in the database.
	\item
	\textbf{Valid time}: The time when the fact is true in the modelled reality.
	\item
	\textbf{Decision time} \cite{Nascimento95}: The time when an event was decided to happen. 
	\end{itemize}

	 The  database model can be then classified into: transaction time \cite{Jensen91}, Valid time, bi-temporal \cite{Snodgrass84}(both valid and transaction time) or tri-temporal \cite{Nascimento95} (valid, transaction and decision time).

Fuzzy temporal models \cite{schockaert08} deal with time points \cite{Dubois89} or intervals \cite{Garrido2009} that are ill-known. Some fuzzy temporal models assume that the time stored in the database is an interval. The temporal interval is represented by two ill-known time points: $X$  an ill-known starting point and $Y$ an ill-known ending point. The interval $\left[X,Y\right]$ is not a fuzzy interval but an ill-known interval: it is a crisp interval but it is partially unknown which values are in this interval.

>>>>

\subsection{\label{subsec:possibilistic-variables}Possibilistic Variables}
A \emph{possibilistic variable} is defined as follows\cite{Pon11}.

\begin{definition}
A possibilistic variable $X$ over a universe $U$ is defined as a variable taking exactly one value in $U$, but for which this value is (partially) unknown. Its possibility distribution $\pi_X$ on $U$ gives the available knowledge about the value that $X$ takes: for each $u\in U$, $\pi_X(u)$ represents the possibility that $X$ takes the value $u$.
\end{definition}

@@

When a possibilistic variable is defined on the powerset $\Pow(R)$ of some universe $R$, the unique value the variable takes will be a crisp set and its possibility distribution on the powerset $\Pow(R)$ will describe the possibility of each crisp subset of $R$ to be the value the variable takes. This exact value (a crisp set) the variable takes, is now called an \emph{ill-known set}\cite{Dubois88b}.

@@

A specific application of possibilistic variables is obtained when the universe under consideration is the set of boolean values $\mathbb{B}$ = $\{T,F\}$. Indeed, any boolean proposition $p$ takes just one value in $\mathbb{B}$. If the knowledge on which value this proposition $p$ will take is given by a possibility distribution $\pi_p$, the proposition can be seen as a possibilistic variable. As the interest lies with the case where the proposition holds, the possibility and necessity that $p$ = $T$ (the proposition holds) demand most attention. This possibility and necessity is noted here as:
\begin{align}
\text{Possibility that $p$ = $T$ (p holds):} \hspace{50pt} & Pos(p) = \pi_p(T) \label{propholdsposs} \\
\text{Necessity that $p$ = $T$ (p holds):} \hspace{50pt} & Nec(p) = 1-\pi_p(F) \label{propholdsnecc}
\end{align}
%Here, notation \ref{propholdsposs} denotes the possibility that $p$ = $T$ and the proposition holds, notation \ref{propholdsnecc} denotes the necessity that $p$ = $T$ and the proposition holds.

\subsection{\label{subsec:fuzzy-numbers}Fuzzy Numbers and Fuzzy Intervals}
Among others, Dubois and Prade~\cite{Dubois1983} use the following definition of a \emph{fuzzy interval}.
\begin{definition}
A fuzzy interval is a fuzzy set $M$, defined by a membership function $\mu_{M}$, on the set of real numbers $\mathbb{R}$ such that:
\begin{eqnarray}
\mu_{M} : & \!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\! \mathbb{R} \rightarrow \left[0,1\right] \nonumber \\ 
\forall (u,v)\in\mathbb{R}^2: \forall w \in [u,v]:&\mu_M(w) \geq\min(\mu_M(u),\mu_M(v))  \\
\exists m \in \mathbb{R} : & \!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\! \mu_M(m)=1 
\end{eqnarray}
\end{definition}
If this modal value $m$ is unique, then $M$ is referred to as a \emph{fuzzy number}. In other words, if the core of a fuzzy interval is a singleton, it is referred to as a fuzzy number.

A simple form of the membership function of a fuzzy interval is a trapezoidal function. It can be shown that such a membership function $\mu_T$ for a fuzzy interval $T$ is convex and normalized. A visualization is given in figure \ref{fig:trapezoidal}. Four reel values, denoted $\alpha$, $\beta$, $\gamma$ and $\delta$ and chosen as in figure \ref{fig:trapezoidal}, suffice to represent a trapezoidal membership function of a fuzzy interval. In this work, a fuzzy interval defined as such will be noted as $\left[\alpha, \beta, \gamma, \delta\right]$. %The corresponding membership function definition for this $\mu_T$ is then given by:

%\begin{align}
%\mu_T : & \quad \mathbb{R} \rightarrow \left[0,1\right] \\
% : & \quad x \rightarrow
%\begin{cases}
%1 & \mbox{ if } x \in [\beta,\gamma] \\
%0 & \mbox{ if } x > \delta \vee x < \alpha \\
%\frac{x-\alpha}{\beta - \alpha} & \mbox{ if } x \in [\alpha,\beta[ \\
%\frac{\delta -x}{\delta - \gamma} & \mbox{ if } x \in ]\gamma,\delta] \\
%\end{cases}
%\end{align}

@@

The most convenient form of the membership function of a fuzzy number is a triangular function. It can be shown that such a membership function $\mu_M$ for a fuzzy number $M$ is convex and normalized. A visualization is given in figure \ref{fig:triangular}. Three reel values, denoted $a$, $b$ and $D$, suffice to represent a triangular membership function of a fuzzy number and in this work, a fuzzy number defined as such will be noted as $\left[D, a, b \right]$. Here:
\begin{itemize}
\item
$D$ denotes the single value in the core of $M$
\item
$D-a$ is then $\inf \{u \in \mathbb{R} : \mu_{M}(u) > 0\}$
\item
$D+b$ is then $\sup \{u \in \mathbb{R} : \mu_{M}(u) > 0\}$
\end{itemize}
\begin{figure}[h!]
  \centering
  \input{./graphs/triangular.tex}
  \caption{Triangular membership function.}
  \label{fig:triangular}
\end{figure}

@@

\subsubsection{Interval Evaluation by Ill-known Constraints}
The problem of interval evaluation is more generally explained in \cite{Pon11}: the need exists to know if all points in a crisp interval $I$ reside between the boundaries of an ill-known interval $\left[ X , Y \right]$. In \cite{Pon11}, the notion of an \emph{ill-known constraint} is introduced:

\begin{definition}
Given a universe $U$, an \emph{ill-known constraint} $C$ on a set $A \subseteq U$ is specified by means of a binary relation $R \subseteq U^{2}$ and a fixed, ill-known value denoted by its possibilistic variable $V$ over $U$, i.e.:
\begin{align}
C \triangleq (R,V)
\end{align}
Set $A$ now satisfies the constraint if and only if:
\begin{align}
\forall a \in A : (a,V) \in R
\end{align}
\end{definition}

The satisfaction of a constraint $C \triangleq (R,V)$ by a set $A$ is now still a Boolean matter, but due to the uncertainty about the ill-known value $V$, it can be uncertain whether $C$ is satisfied by $A$ or not\cite{Pon11}. In fact, this satisfaction now behaves as a proposition. Based on the possibility distribution $\pi_{V}$ of $V$, the possibility and necessity that $A$ satisfies $C$ can be found. This proposition can thus be seen as a possibilistic variable on $\mathbb{B}$. The required possibility and necessity are:

\begin{align}
\Pos(A\text{ satisfies }C) & = \min_{a \in A}\left(\sup_{(a,w) \in R}\pi_{V}(w))\right) \\
\Nec(A\text{ satisfies }C) & = \min_{a \in A}\left(\inf_{(a,w) \notin R} 1-\pi_{V}(w)\right)
\end{align}

Now, to check if crisp interval $I = \left[j, k\right]$ is included in $\left[X, Y\right]$, 2 ill-known constraints are constructed:

%We assume that $X$ specifies the lower bound and $Y$ the upper bound for a given interval, we want to known whether all points in the interval are larger than or equal to $X$ and smaller than or equal to $Y$. Therefore, we consider two ill-known constraints:

\begin{eqnarray}
C_1\triangleq\left(\geq,X\right)\\
C_2\triangleq\left(\leq,Y\right).
\end{eqnarray}

The possibility and necessity of the set satisfying both constraints is then:

\begin{align}
\label{eq:interval-pos}
\Pos(A\text{ satisfies }C_1\ AND\ C_2) & = \min_{a \in A}\left(\sup_{a \geq w}\pi_{X}(w),\sup_{a \leq v}\pi_{Y}(v)\right)\\
\label{eq:interval-nec}
\Nec(A\text{ satisfies }C_1\ AND\ C_2) & = \min_{a \in A}\left(\inf_{a < w} 1-\pi_{X}(w),\inf_{a > v} 1-\pi_{Y}(v)\right).
\end{align}

@@

\subsection{Temporal Databases}
%In this section, the proposed reasoning is applied to the specific context of intervals on the real line. This setting is of specific interest in the context of fuzzy temporal databases.
A \emph{temporal database} is a database that manages some aspects of time in its schema. The reality a temporal database tries to model, contains some temporal notions which have to be handled specifically in order to maintain a consistent modelling behavior. A \emph{chronon} is the shortest duration of time supported by the database. Time can be represented either as points or intervals \cite{655777}.
%There are proposals for the fuzzyfication of the time point and the fuzzyfication of the time interval.

@@

The temporal notions in temporal databases can be classified into 4 types based on their interpretation and modelling purpose. User-defined time has no interpretation, but the other types do:

\begin{itemize}
	\item
	\emph{Transaction time}: The time when the fact is stored in the database.
	\item
	\emph{Valid time}: The time when the fact is true in the modelled reality.
	\item
	\emph{Decision time}: The time when an event was decided to happen. 
\end{itemize}

Database models can then be classified into transaction time \cite{Jensen91}, decision time, valid time, bi-temporal \cite{Snodgrass84}(both valid and transaction time) or tri-temporal \cite{Nascimento95} (valid, transaction and decision time) models.

To deal with time points \cite{Dubois89} or intervals \cite{Garrido2009} that are subject to imperfection and can thus not be precisely known, fuzzy temporal models \cite{schockaert08} exist. %Some fuzzy temporal models assume that the time stored in the database is an interval. The temporal interval is represented by two ill-known time points: $X$  an ill-known starting point and $Y$ an ill-known ending point. The interval $\left[X,Y\right]$ is not a fuzzy interval but an ill-known interval: it is a crisp interval but it is partially unknown which values are in this interval.
